### 2019-01-01 index.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/index-",
  out.width = "100%"
)
```

# areal <img src="man/figures/arealLogo.png" align="right" />

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis-CI Build Status](https://travis-ci.org/slu-openGIS/areal.svg?branch=master)](https://travis-ci.org/slu-openGIS/areal)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/slu-openGIS/areal?branch=master&svg=true)](https://ci.appveyor.com/project/chris-prener/areal)
[![Coverage status](https://codecov.io/gh/slu-openGIS/areal/branch/master/graph/badge.svg)](https://codecov.io/github/slu-openGIS/areal?branch=master)
[![CRAN_status_badge](http://www.r-pkg.org/badges/version/areal)](https://cran.r-project.org/package=areal)

Areal interpolation is the process making estimates from a source set of polygons to an overlapping but incongruent set of target polygons. One challenge with areal interpolation is that, while the processes themselves are well documented in the academic literature, implementing them often involves "reinventing the wheel." While the `R` package `sf` does offer a basic interface for areal weighted interpolation (`st_interpolate_aw`), it lacks some features that we use in our work. The `areal` package contains a suite tools for validation and estimation, providing a full-featured workflow that fits into both modern data management (e.g. `tidyverse`) and spatial data (e.g. `sf`) frameworks.

## Bug Fixes in v0.1.3
The initial CRAN release contains two known bugs (see Issues [6](https://github.com/slu-openGIS/areal/issues/6) and [7](https://github.com/slu-openGIS/areal/issues/7)), both of which are fixed in the current GitHub release:

1. If the `tid` and `sid` column names are identical, the interpolated column created by `aw_interpolate()` will consist of all `NA` values. The simplest workaround is to rename either `tid` or `sid` before executing `aw_interpolate()`. Alternatively, you can install the development version of `areal` from GitHub.
2. Some intersections of `sf` geometries will result in geometry collections being created. This appears to be an infrequent issue, but it will result in an error when `aw_interpolate()` is executed. The error will begin with `Warning in st_cast.GEOMETRYCOLLECTION(X[[i]], ...)`. The simplest workaround is to install the development version of `areal` from GitHub.

Big thanks to early adopter [Matt Herman](https://github.com/mfherman) for catching both of these!

## Quick Start
If the `sf` package is already installed, the easiest way to get `areal` is to install it from CRAN:

``` r
install.packages("areal")
```

Alternatively, the development version of `areal` can be accessed from GitHub with `remotes`:

```r
# install.packages("remotes")
remotes::install_github("slu-openGIS/areal")
```

Additional details, including some tips for installing `sf`, can be found in the [Get started article](articles/areal.html#getting-started).

## Resources
In addition to instructions for installation, the main [Get started](articles/areal.html) article has:

  * a quick overview of areal interpolation,
  * some notes on preparing data for interpolation,
  * a brief introduction to the `aw_interpolate()` function,
  * tips for getting help and submitting feedback, 
  * and the `areal` package's development roadmap!

This site also offers dedicated articles on [data preparation](articles/data-preparation.html) and [using `areal` for areal weighted interpolation](articles/areal-weighted-interpolation.html).